---
title: "Percent vs fold change"
author: "Renan EC"
date: "8/25/2020"
output: html_document
---

# Quick notes on comparing fold change versus percentage

A fold change FC is defined as 

$$FC = \frac{y_t}{y_0}$$
where $y_t$ is the value of the variable y at time t and $y_0$ is the variable at time zero.

We can express the  change of $y_t$ with respect to $y_0$ as:

$$change = y_t-y_0$$
And when defined as percentages we get:

$$
\begin{aligned}
y_0 \leftrightarrow 100\% \\
y_t \leftrightarrow \frac{y_t}{y_0}*100 \%\\
\end{aligned}
$$
So then we can express $y_t - y_0$  as a function of percents

$$
\begin{aligned}
y_t - y_0 = \frac{y_t}{y_0}*100 - 100 \\
y_t - y_0 = 100 \times [ \frac{y_t}{y_0} - 1] \\
y_t - y_0 = 100 \times [ FC - 1] \\
percChange = 100 \times [ FC - 1] \\
\end{aligned}
$$
$$
percChange = 100 \times [ FC * - 1] \\
FC = \frac{percChange}{100} + 1
$$

Does this idea make sense

What is a 4 fold change difference in percentage?

```{r compute_fold_change}

compute_fold_change <- function(yt, y0){
  
  FC <- yt/y0
  return(FC)
}

```

What is the mapping between FC and % change?

```{r}
compute_percent_change <- function(yt, y0){
  
  PC <- (yt/y0 - 1) * 100
  
  return(PC)
  
}
```


```{r}
compute_fold_change(100, 25)
compute_percent_change(100, 25)
```

```{r}

```

