---
title: "Percent vs fold change"
author: "Renan EC"
date: "8/25/2020"
output: html_document
---

```{r, include=FALS}
library(tidyverse)
library(magrittr)
library(knitr)
library(scales)
library(plotly)
```

# Quick notes on comparing fold change versus percentage

A fold change FC is defined as 

$$FC = \frac{y_t}{y_0}$$
where $y_t$ is the value of the variable y at time t and $y_0$ is the variable at time zero.

We can express the  change of $y_t$ with respect to $y_0$ as:

$$change = y_t-y_0$$
And when defined as percentages we get:

$$
\begin{aligned}
y_0 \leftrightarrow 100\% \\
y_t \leftrightarrow \frac{y_t}{y_0}*100 \%\\
\end{aligned}
$$
So then we can express $y_t - y_0$  as a function of percents

$$
\begin{aligned}
y_t - y_0 = \frac{y_t}{y_0}*100 - 100 \\
y_t - y_0 = 100 \times [ \frac{y_t}{y_0} - 1] \\
y_t - y_0 = 100 \times [ FC - 1] \\
percChange = 100 \times [ FC - 1] \\
\end{aligned}
$$
$$
percChange = 100 \times [ FC * - 1] \\
FC = \frac{percChange}{100} + 1
$$

Does this idea make sense

What is a 4 fold change difference in percentage?

```{r compute_fold_change}

compute_fold_change <- function(yt, y0){
  
  FC <- yt/y0
  return(FC)
}

```

What is the mapping between FC and % change?

```{r}
compute_percent_change <- function(yt, y0){
  
  PC <- (yt/y0 - 1) * 100
  
  return(PC)
  
}
```


```{r}
compute_fold_change(100, 25)
compute_percent_change(100, 25)
```

## Let's create a table with values and display percent changes

```{r}
plot_data_df <- tibble(
  y_t = seq(from = -1000, to = 1000, by = 100)
)


plot_data_df %<>% 
  mutate(y_0 = 100) %>% 
  mutate(FC = compute_fold_change(y_t, y_0),
         PC = compute_percent_change(y_t, y_0)) 
```

```{r}
plot_data_df %>% 
  kable()
```


```{r}
plot_data_df %>% 
  ggplot(aes(x=FC, y= PC)) +
  geom_point() 
```
## Interconverting between FC and PC

Now let's learn how to interconvert between the two of them with the library scales

```{r}
p <- plot_data_df %>% 
  pivot_longer(cols = c("FC", "PC"), names_to = "change") %>% 
  ggplot(aes(x=y_t , y = value, color = change)) +
  geom_point()

ggplotly(p)
```

```{r}
 plot_data_df %>% 
  ggplot(aes(x=y_t, y=PC)) +
  geom_point()
```

We need to define an object of type trans. We need to define a transform and inverse for labeling to work properly


```{r}
FC_to_PC <- function(x) (x - 1) * 100
PC_to_FC <- function(x) (x/100) + 1

percent_to_fold_change_trans <- function() trans_new(name = "percent_to_fold_change", transform = PC_to_FC, inverse = FC_to_PC)
```

```{r}
plot_data_df %>% 
  mutate(FC_new = PC_to_FC(PC),
         difference = FC_new - FC) %>% 
  pull("difference") %>% 
  sum()
```

```{r}
p <- plot_data_df %>% 
  ggplot(aes(x=y_t, y=PC)) +
  geom_point() +
  coord_trans(y = "percent_to_fold_change")
p
```




```{r}
one_over_trans = function() trans_new(name = "one_over", transform = function(x) 1/x, inverse = function(x) x)
```

```{r}
qplot(carat, price, data=diamonds) +
  coord_trans(y = "one_over")

# + coord_trans(x = "log10", y = "one_over")
```

