---
title: "Understanding gr metric"
output: html_document
---


Understanding gr

```{r}
library(tidyverse)
library(heatmaply)
library(plotly)
```


## Definition of gr


Is it only defined for positive values?

$$gr= 2^\frac{gr_{treated}}{gr_{untreated}}-1$$

```{r}
compute_gr <- function(gr_drug_treated, gr_untreated){
  
  gr <- 2^(gr_drug_treated/gr_untreated) - 1 
  return(gr)
}
```

1 means no inhibition (the gr drug treated and untreated are the same)

```{r}
compute_gr(gr_drug_treated = -30000, gr_untreated = 1)
```

Partial growth inhibition

```{r}
compute_gr(gr_drug_treated = 1, gr_untreated = 2)
```

When the treatment leads to no growth then the gr metric is zero

```{r}
compute_gr(gr_drug_treated = 0, gr_untreated = 2)
```
When there is a negative gr with the drug the gr is negative
When cells are dying at the same rate they used to divide then the gr metric is -0.5
```{r}
compute_gr(gr_drug_treated = -2, gr_untreated = 2)
```

What if there is stimulation of growth, then it's above 1p

```{r}
compute_gr(gr_drug_treated = 0, gr_untreated = 2)
```

## Plot curve

```{r}
f1 <- function(x){
  x^2
}

curve(f1, from = -10, to= 10)
```
```{r}
plot(f1, -10, 10)
```
```{r}
ggplot(data.frame(x=c(-10,10)), aes(x=x)) +
  stat_function(fun = f1)
```


## Computing 2D functions

https://stackoverflow.com/questions/47879000/how-to-graph-function-with-two-variables-in-r

```{r}
gr_treated <- seq(-3, 3, by=0.1)
gr_untreated <- c(-3, -2, -1, 1, 2, 3)

computed_gr_mat <- mapply(compute_gr, list(gr_treated), gr_untreated)
```

###  Testing perspective plots

https://stackoverflow.com/questions/22150235/multidimensional-2d-function-plot-in-r

```{r}
persp(gr_treated,gr_untreated,computed_gr_mat, col='blue')
```

```{r}
par(mar=c(1,1,1,1))
layout(matrix(1:4, nrow=2))
s=lapply(c(0,30,60,90), function(t) persp(gr_treated,gr_untreated,computed_gr_mat, col='blue', theta=t))
```


```{r}
cols <- RColorBrewer::brewer.pal(8, "Accent")
matplot(gr_treated, computed_gr_mat, col=cols, type="l", lty=1, lwd=2, xlab="gr_treated", ylab="result")
legend("bottomright", legend=gr_untreated, title="gr_untreated", lwd=2, col=cols)
```
```{r}
image(computed_gr_mat)
```

$$gr= 2^\frac{gr_{treated}}{gr_{untreated}}-1$$
## Visualize gr metric using ggplot

```{r}
res_df <- computed_gr_mat %>% as.data.frame()
colnames(res_df) <- str_c("gr_untreated", gr_untreated, sep="#")
rownames(res_df) <- str_c("gr_treated", gr_treated, sep = "#")

tidy_gr_df <- res_df %>% 
  tibble::rownames_to_column(var = "gr_treated") %>% 
  as_tibble() %>% 
  pivot_longer(cols= -gr_treated ,names_to = "gr_untreated", values_to = "gr") %>% 
  mutate(gr_treated = str_remove(gr_treated, "gr_treated#"),
         gr_untreated = str_remove(gr_untreated, "gr_untreated#"))
```

```{r, fig.width=3, fig.height=3}
point_plot <- tidy_gr_df %>% 
  ggplot(aes(x = gr_untreated, y= gr_treated, size= gr, color= gr)) +
  geom_point()
```

```{r}
plotly::ggplotly(point_plot)
```

```{r,  fig.width=3, fig.height=3}
 # tidy_gr_df %>% 
 #  ggplot(aes(x = gr_untreated, y= gr_treated, z= gr)) +
 #  geom_tile()
```

```{r,  fig.width=3, fig.height=3}
ggplot(tidy_gr_df, aes(gr_untreated, gr_treated)) +
 geom_raster(aes(fill = gr)) +
  scale_fill_gradientn(colours=c("yellow","red"))
```

```{r}
tidy_gr_df %>% 
  mutate(gr_untreated = as.factor(gr_untreated)) %>% 
  ggplot(aes(x=gr_treated, y = gr, color = gr_untreated)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = -1) +
  # geom_vline(xintercept = 0, color = "red") +
  # geom_vline(xintercept = 1, color = "red") +
  geom_point() +
  # geom_vline(xintercept = 10)
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) +
  theme_bw()
  
```

```{r}
fig <- plot_ly(
    x = gr_treated, 
    y = gr_untreated,
    z = res, type = "heatmap",
    color
)

fig
```
I would like 1 to be a color, 0 a different distinct color as well as -1. Then gradations from there work
```{r}
res_df <- res %>% as.data.frame()
colnames(res_df) <- str_c("gr_untreated", gr_untreated, sep="#")
rownames(res_df) <- str_c("gr_treated", gr_treated, sep = "#")
```


```{r}
heatmaply(res_df)
```

